Novel Methods for Characterizing Peaks from Direct Injection FT-MS Experiments

### 2019-12-24 peakcharacterization_manuscript.Rmd
discuss hpd sites

For each raw data file in mzML format, they were each processed in these six ways:

  1. No noise removal, no normalization, no frequency SD filtering
  2. Noise removal, no normalization, no frequency SD filtering
  3. Noise removal, single pass normalization with all peaks, no frequency SD filtering
  4. Noise removal, single pass normalization with high ratio peaks, no frequency SD filtering
  5. Noise removal, two pass normalization, no frequency SD filtering
  6. Noise removal, two pass normalization, frequency SD filtering

@@

>>>>

For each raw data file in profile mzML format, they were processed in these six ways:

  1. No noise removal, no normalization, no frequency SD filtering (noperc_nonorm)
  2. Noise removal, no normalization, no frequency SD filtering (perc99_nonorm)
  3. Noise removal, single pass normalization with all peaks, no frequency SD filtering (singlenorm)
  4. Noise removal, single pass normalization with high ratio peaks, no frequency SD filtering (singlenorm_int)
  5. Noise removal, two pass normalization, no frequency SD filtering (doublenorm)
  6. Noise removal, two pass normalization, frequency SD filtering (filtersd)

@@

### Removal of High Peak Density Artefacts

We previously described the presence of high peak density (HPD) artefacts in FT-MS spectra [Mitchell et al., Metabolomics, 2018]. 
Ideally, the peak characterization procedure should reduce their presence in the resultant reported peaks. 
Their presence should be minimized by removing noise peaks, and removing regions that have multiple reported peaks in the same scan. 
We expect they may present are as characterized peaks that have larger than expected standard deviations in their frequencies when calculated across scans.
These peaks can be detected by simply examining the distribution of frequency SDs and removing those that are outliers.
To verify the removal of HPD regions, we converted centroided M/Z's from XCalibur to frequency values using the previously decided values for that sample.
Peak density was measured using a sliding window 10 points wide and offset by one point.
